DeepSphere: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications

### 2019-03-04 rebuttal.tex

\section*{Answer to the editor}

\begin{mdframed}[style=comment]
Thank you for submitting your manuscript to Astronomy and Computing. I have received comments from reviewers on your manuscript. Your paper should become acceptable for publication pending suitable minor revision and modification of the article in light of the appended reviewer comments.  In particular, there is a request from a reviewer to have access to the data used in your analysis.

Though this is designated a minor revision, there are quite a few points that need to be addressed before the paper is acceptable for publication.  When resubmitting your manuscript, please carefully consider all issues mentioned in the reviewers' comments, outline every change made point by point, and provide suitable rebuttals for any comments not addressed.
\end{mdframed}

We thank the editor for handling our submission and organizing the peer-review process.
We have revised our manuscript following the suggestions of the reviewers.
We believe that we addressed most requests and that the manuscript is now ready for publication.
We thank the reviewers for taking the time to review our manuscript and provide constructive comments that improved the quality of our contribution.

Currently, access to the dataset can be requested via Zenodo.
As this process is not anonymous, we provide in this rebuttal (comment \#2) a link for the reviewers to access it anonymously.

@@

% \mdeff{Agreed that we should do at least one thing they asked for.
% \begin{itemize}
% 	\item Testing on the 1/12 of the sphere is not what he asked for. On one hand, a global projection would exhibit more distortion. On the other, we never use the whole sphere.
% 	\item Keep in mind that there is no orientation that is consistent for the 12 base faces (i.e., where is up?).
% 	\item Agreed that the performance should be very similar. Both are doing local computations. And locally (3 to 5 pixels across), the sphere is isomorphic to an Euclidean plane.
% 	\item The only difference, in my opinion, would be due to the anisotropic filters the 2D CNN has access to.
% 	\item Not sure we should include this result in the main paper. Hard to justify in the story, especially as a comparison with alternative spherical CNNs is missing. Let's see the results.
% \end{itemize}
% }

\subsection{}

\begin{mdframed}[style=comment]
Similarly, I think it would useful to check how well the DeepSphere CNN does against the FCN when the input data is rotated randomly during training - I imagine that the results should match or probably exceed the FCN results and I think this would be a neat little test which doesn't involve needing extra data, just more augmentation.
\end{mdframed}

% That comparison is not warranted.
% or justified. pointless is maybe too aggressive
We do not believe this comparison is useful because of the following reason.
There are two ways to deal with symmetries one wants to be invariant to: (i) build them into the architecture of the model, or (ii) augment the dataset such that a (more general) model learns them ``brute-force''.
With respect to rotation invariance, the FCN architecture is of the first kind, while the CNN is of the second kind.
Assuming the task is truly invariant to said symmetry,\footnote{As argued in the introduction, cosmological maps are rotation invariant or equivariant.} the performance of the first kind of models would be exactly the same as without augmentation, while the second kind would only catch up after seeing enough data to learn the invariance.
% That is a waste of resources which is only justified if it's not possible to back the desired invariance in the architecture.
Augmentation is only justified if it's not possible to back the desired invariance in the architecture.
As the reviewer writes in comment 2.8, we should be ``building a NN to suit [our] purposes''.
We updated the manuscript to more precisely reflect this argumentation.

### 2019-03-05 scnn.tex
update the figures

\begin{figure}[t!]
    \centering
    \includegraphics[width=\linewidth]{filtering_speed}
    \caption{Comparison of filtering speed for Gaussian smoothing of maps of various sizes.
    The fast spherical harmonic transform (SHT) is implemented by the \texttt{healpy} Python package (via the \texttt{healpy.sphtfunc.smoothing} function).
    The graph filtering is defined by \eqnref{graph_convolution_cheby} and implemented with the \texttt{numpy} and \texttt{scipy} Python packages.
    Both are executed on a single core.
	The theoretical cost of filtering on the graph is $\bO(K N_{pix})$ and $\bO(\ell_{max}^3) = \bO(N_{pix}^{3/2})$ for the spherical harmonics, where $\ell_{max}$ is the largest angular frequency.}
    \label{fig:filtering_speed}
\end{figure}

@@

Exploiting the recursive formulation of Chebyshev polynomials, evaluating \eqnref{graph_convolution_cheby} requires $\bO(K)$ multiplications of the vector $\f$ with the sparse matrix $\tL$.
The cost of one such multiplication is $\bO(|\E| + |\V|)$.
By construction of our graph, $|\E| < 8 N_{pix}$ and the overall computational cost of the convolution reduces to $\bO(N_{pix})$ operations and as such is much more efficient than filtering with spherical harmonics, even though HEALPix was designed as an iso-latitude sampling that has a fast spherical transform.
This is especially true for smooth kernels which require a low polynomial degree $K$.
\figref{filtering_speed} compares the speed of low-pass filtering for Gaussian smoothing using the spherical harmonics and the graph-based method presented here.
On a single core, a naive implementation of our method is ten to twenty times faster for $N_{side} = 2048$, with $K=20$ and $K=5$, respectively, than using the spherical harmonic transform at $\ell_{max} = 3 N_{side}$ implemented by the highly optimized \pkg{libsharp} \citep{reinecke2013libsharp} library used by HEALPix.
Note that the two might however scale differently, given that \pkg{libsharp} can be distributed on CPUs through MPI while our method can be distributed on GPUs by \pkg{TensorFlow}.

>>>>

\begin{figure}[t!]
    \centering
    \includegraphics[width=\linewidth]{filtering_speed}
    \caption{Comparison of filtering speed for Gaussian smoothing of maps of various sizes.
    The fast spherical harmonic transform (SHT) is implemented by the \texttt{healpy} Python package (via the \texttt{healpy.sphtfunc.smoothing} function).
    The graph filtering is defined by \eqnref{graph_convolution_cheby} and implemented with the \texttt{numpy} and \texttt{scipy} Python packages.
    Both are executed on a single core.
	The theoretical cost of filtering on the graph is $\bO(K N_{pix})$ and $\bO(\ell_{max}^3) = \bO(N_{pix}^{3/2})$ for the spherical harmonics, where $\ell_{max}$ is the largest angular frequency.
	The timings for the partial graphs correspond to the convolution on only a fraction (1/12, 1/48, 1/192) of the sphere. This illustrates how graph convolution scales with $N_{pix}$.
	}
    \label{fig:filtering_speed}
\end{figure}

@@

Exploiting the recursive formulation of Chebyshev polynomials, evaluating \eqnref{graph_convolution_cheby} requires $\bO(K)$ multiplications of the vector $\f$ with the sparse matrix $\tL$.
The cost of one such multiplication is $\bO(|\E| + |\V|)$.
By construction of our graph, $|\E| < 8 N_{pix}$ and the overall computational cost of the convolution reduces to $\bO(N_{pix})$ operations and as such is much more efficient than filtering with spherical harmonics, even though HEALPix was designed as an iso-latitude sampling that has a fast spherical transform.
This is especially true for smooth kernels which require a low polynomial degree $K$.
\figref{filtering_speed} compares the speed of low-pass filtering for Gaussian smoothing using the spherical harmonics and the graph-based method presented here.
On a single core, a naive implementation of our method is ten to twenty times faster for $N_{side} = 2048$, with $K=20$ and $K=5$, respectively, than using the spherical harmonic transform at $\ell_{max} = 3 N_{side}$ implemented by the highly optimized \pkg{libsharp} \citep{reinecke2013libsharp} library used by HEALPix.
We also observe an important gain for graph convolution if it only needs to process only a subpart of the sphere. 
This gain is the direct reflection of the graph convolution complexity proportional to $N_{pix}$.
Note that the two might however scale differently, given that \pkg{libsharp} can be distributed on CPUs through MPI while our method can be distributed on GPUs by \pkg{TensorFlow}.
